Prestige as a Driving Force in Cultural Transmission

### 2022-03-28 ms/PrestigeBias.tex
revise abstract

\begin{abstract}
Copying role-models can be an efficient method of acquiring knowledge. Here, we study the effects of prestige as a cultural transmission bias in cultural-evolution model. A common bias when choosing a role-model to copy is success bias, i.e copying whoever appears more successful. This bias depends on the performance of the role-model alone, without other factors. We propose an additional bias that may be prevalent in cultural transmission: influence bias, which is evaluated by the number of individuals that have already copied a specific role-model. We combine these biases together to a prestige bias and analyze its effects on the evolutionary dynamics of the population using mathematical analysis and stochastic simulations. We find analytic approximations to our model, facilitating further mathematical analysis and reducing the computational complexity of the simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
For a dichotomous cultural trait, we find  an approximation to the fixation probability and the fixation time of an invading advantageous cultural trait, in both a constant and changing environment, which resembles Kimura's classical formulas.
These approximations show that influence effectively reduces the population size. % TODO reduce or increase?
We also find that influence accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model may provide a better description of how human cultural transmission, 
especially in the last years where social networks are very popular. % TODO to discuss this last sentence
Further work is needed to test if this model could predict various phenomena in human cultural evolution when extended with the effects of selection and innovation.
\end{abstract}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

>>>>

\begin{abstract}
Copying role-models can be an efficient method for acquiring knowledge. A common bias when choosing a role-model to copy is success bias: copying whoever appears more successful. This bias depends on the performance of the role-model alone, with no other factors. We propose an additional bias that may be prevalent in cultural transmission: influence bias, in which role-model choice is affected by the number of individuals that have already copied each potential role-model. We combine success and influence bias together to a ``prestige bias'' and analyze its effects on cultural-evolutionary dynamics using mathematical analysis and stochastic simulations. We find analytic approximations to our stochastic model, facilitating further mathematical analysis and reducing the computational complexity of simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
We also find approximations to the fixation probability and the fixation time of an invading advantageous cultural trait, in both a constant and changing environment, which resemble Kimura's classical formulas.
These approximations show that success bias effectively plays the part of natural selection, whereas influence bias effectively reduces the population size. % TODO reduce or increase?
It also accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model may provide a better description of human cultural transmission, 
especially in the last years where social networks are very popular. % TODO to discuss this last sentence
Further work is needed to test if this model could predict various phenomena in human cultural evolution when extended with the effects of selection and innovation.
\end{abstract}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}
